---
title: |
  | Chapter 4
  | Elevated atmospheric CO~2~ and drought does not alter total belowground carbon allocation in *Eucalyptus saligna*
author: ""
date: ""
output:
  word_document:
    reference_docx: manuscriptstyle_thesis.docx
  pdf_document: default
  html_document:
    number_sections: yes
csl: tree-physiology.csl
bibliography: references.bib
---

<!-- \ -->
<!-- Courtney E. Campany^1^, Craig Barton^1^, Ross E. McMurtrie^2^, Remko A. Duursma^1^. -->

<!-- \ -->
<!-- ^1^ Hawkesbury Institute for the Environment, University of Western Sydney, Locked Bag 1797, Penrith, NSW, Australia -->

<!-- \ -->
<!-- ^2^ School of Biological, Earth and Environmental Sciences, University of New South Wales, Sydney, NSW, Australia -->

<!-- \ -->
<!-- *Corresponding author*:  -->
<!-- Courtney Campany -->
<!-- E: courtneycampany@gmail.com -->


```{r, results='asis', echo=FALSE, warning=FALSE}
library(knitr)
opts_knit$set(root.dir = '../')
```


```{r global sourcing , include=FALSE}
source("functions_and_packages/plot_objects.R")
source("functions_and_packages/functions.R")
library(plotrix)
library(doBy)

##table data
library(pixiedust)
table <- read.csv("master_scripts/data_table.csv", stringsAsFactors = FALSE)

Pinter <- read.csv("Stats/p_sigs/P_interactions.csv")
Pinter <- round(Pinter, 3)

Pco2 <- read.csv("Stats/p_sigs/P_co2.csv")
Pco2 <- round(Pco2, 3)

Ph2o <- read.csv("Stats/p_sigs/P_h20.csv")
Ph2o <- round(Ph2o, 3)

###sla P values
Pinter_sla <- read.csv("Stats/p_sigs/sla/P_interactions_sla.csv")
Pinter_sla <- round(Pinter_sla, 3)

Pco2_sla <- read.csv("Stats/p_sigs/sla/P_co2_sla.csv")
Pco2_sla <- round(Pco2_sla, 3)

Ph2o_sla <- read.csv("Stats/p_sigs/sla/P_h20_sla.csv")
Ph2o_sla <- round(Ph2o_sla, 3)

###Add together
Pinter_2 <- rbind(Pinter, Pinter_sla)
Pco2_2 <- rbind(Pco2, Pco2_sla)
Ph2o_2 <- rbind(Ph2o, Ph2o_sla)
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
# Two useful packages for printing numbers, accessing content of objects.
library(broom)
library(reporttools)

# Set Flag to include figures in text or not.
includeFigs <- TRUE

# This document can be compiled with the 'Knit Word' button in Rstudio, or:
# library(rmarkdown)
# render("manuscript.Rmd", "word_document", "manuscript.docx")
```

```{r, echo=FALSE}
conceptfig <- "4.1"
fluxtreefig <-  "4.2"
lafig <- "4.3"
fluxlafig <- "4.4"
massfracfig <- "4.5"
additivemassfig <- "4.6"
belowbarfig <- "4.7"
tbcafluxfig <- "4.8"
rootshootpred <- "4.S1"
additivechamfig <- "4.S2"

wtc1tab1 <- "4.1"
```


#4.1  Abstract
Accurately measuring carbon (C) allocation in large trees above and belowground remains a difficult task and is challenging to represent in models of forest C cycling. Understanding how global change impacts the distribution of tree photosynthate is an essential process in determining future terrestrial C balance. We utilized climate-controlled whole tree chambers (WTCs) to measure cumulative net aboveground CO~2~ uptake of *Eucalyptus saligna* trees, which was expected to correlate to harvested tree C mass. We then investigated how elevated atmospheric CO~2~ concentration and a 4-month drought period affected both tree biomass partitioning and the allocation of photosynthetic C to various above and belowground pools. We calculated total belowground C allocation (TBCA) for each WTC, which includes biomass and all belowground fluxes, as the residual between daily aboveground net CO~2~ uptake and aboveground C mass accrual. It was hypothesized that that both drought and elevated CO~2~ would increase biomass partitioning to roots, as well as TBCA. Cumulative aboveground net CO~2~ uptake correlated positively to both whole tree C mass and mean leaf area over the entire 11 month measured chamber flux period. Surprisingly, biomass partitioning to roots and cumulative TBCA were unaffected by either elevated CO~2~ or drought. As a fraction of total aboveground net C flux, TBCA remained relatively stable (ca. 40%) across the final 11 months of the experiment for all trees. Carbon allocation to leaves increased under elevated CO~2~, while the effects of a 4 month drought were negligible on biomass production or C allocation of aboveground tissues. The novel approaches used here provide evidence that belowground processes may not be as sensitive to global change as previously thought. These results reveal how quantifying the investment of photosynthetic C beyond biomass production is key to assessing functional tree growth responses, while also providing an empirical framework to test model representations of C allocation in trees. 

<!-- ##Key Words -->
<!-- carbon allocation, biomass partitioning, *Eucalyptus*, elevated CO~2~, drought -->

#4.2  Introduction
Carbon (C) allocation in trees encompasses investment into biomass production above and belowground as well as fluxes including tissue respiration and exudation [@litton2007carbon]. Trees must allocate C to maximize competitive fitness, reproduction and growth across their life cycle [@dickson1989carbon]. In resource-saturated environments plants should maximize growth by allocating resources to support leaf growth to increase C acquisition [@monsi2005factor]. Fluctuations in water, nutrient and light availability, however, may cause plants to invest in roots for belowground resources or stem elongation for increased light harvesting [@friedlingstein1999toward]. These potential changes in C investment are part of a dynamic system:  as the tree grows or sink activities are altered, the fate of C assimilate can shift through time. Understanding allocation is vital, as partitioning among plant organs and their feedback processes profoundly impacts plant growth [@friedlingstein1999toward; @lacointe2000carbon; @shipley2002balanced].

\
Variation in C allocation responses to environmental change combined with a lack of understanding of the mechanisms driving C allocation impede accurate modelling of terrestrial C cycling [@friedlingstein1999toward; @landsberg2003modelling; @litton2007carbon; @epron2012introduction; @mcmurtrie2013new]. The representation of C allocation lags behind leaf photosynthesis (*A~n~*) in process-based forest models [@friedlingstein1999toward; @franklin2012modeling; @iversen2014terrestrial] and this deficiency is due to the difficulty in defining principles that are valid under a wide range of conditions [@franklin2012modeling; @makela2012guiding]. Partitioning coefficients or fixed fractions of assimilation to individual components are often used in models of forest C cycling [@litton2007carbon; @franklin2012modeling]. Unfortunately, using inappropriate or over-simplified allocation schemes can lead to models producing unintended responses or giving the expected answer for the wrong reason [@de2014does;@fatichi2014moving]. As a result, there is continued need to empirically measure patterns of tree C allocation under multi-factor global change manipulations to better understand shifts in future forest C balance.

\
The allocation of photosynthate above and belowground is an important factor in terrestrial C cycling yet our knowledge of how global change drivers impact C allocation is incomplete [@litton2007carbon; @warren2012timing]. With rising atmospheric CO~2~, forest C allocation has drawn particular interest due to its potential effect on C sequestration and the global C balance [@franklin2012modeling]. Across four forested free-air CO~2~ enrichment (FACE) experiments the total flux of C belowground (TBCA), which includes biomass and all belowground fluxes, was found to be enhanced under elevated CO~2~ (eC~a~) [@palmroth2006aboveground]. In FACE experiments this enhancement has been attributed to factors such as increases in C allocation to root biomass production [@iversen2010digging] and root exudation [@phillips2011enhanced]. Alternatively, a meta-analysis by Poorter et al. [-@poorter2000role] concluded that on average, the distribution of biomass to roots, stems or leaves did not change in herbaceous and woody plants grown under eC~a~. 

\
Understanding forest responses to global change also depends on disentangling complex relationships between interacting factors [@rustad2008response]. For example, drought stress in trees can have deleterious effects on leaf [@bradford1982physiological; @schulze1987plant; @broeckx2014effect], stem [@brando2008drought] and root production [@meier2008belowground; @anderegg2012complex]. It has also been shown that C allocation to root systems can increase in drought environments when the severity and duration of the drought periods are substantial [@poorter2012biomass]. The impacts of leaf water savings during CO~2~ enrichment may also enhance tree biomass production under drought conditions [@atwell2007canopy], but sustained enhancement is limited by the availability of a droughted soil water supply to support larger overall canopies [@health1997effects]. The effects of drought may limit C sequestration by the terrestrial biosphere [@iversen2014terrestrial], yet how limitations imposed by drought interact with the growth-stimulating effects of eC~a~ requires more attention [@duursma2011rooting]. 

\
Despite its importance, data on TBCA remain sparse as reliable estimates of root biomass, exudation, turnover and respiration in field conditions are difficult to obtain [@cheng2005measuring; @litton2007carbon; @phillips2008new; @strand2008irreconcilable; @poorter2012biomass]. In forest ecosystems, TBCA has been shown to be equal or greater than aboveground production [@law1999seasonal], yet the controls of this belowground flux are poorly understood [@raich1989belowground; @giardina2005response]. In stand or ecosystem studies, TBCA is often estimated as a residual, by subtracting the changes in C pools of litter, soil and roots from total soil CO~2~ efflux [@raich1989belowground;@davidson2002belowground;@giardina2002total;@palmroth2006aboveground; @adair2009interactive]. A key assumption of this approach is that C pools are in steady-state conditions [@raich1989belowground], which is not always true. Additionally, the reliance on soil respiration in this approach is problematic as studies are often forced to scale up short-term measurements (often monthly) to cumulative yearly fluxes, while also using a variety of measurement techniques. As allocation of C belowground remains one of the most difficult components of tree C budgets to calculate, new approaches are needed to in order accurately track and account for the investment of C belowground.

\
The whole-tree chambers (WTC), located at the Hawkesbury Forest Experiment, were designed to allow continuous measurement of net tree CO~2~ fluxes, allowing canopy *A~n~* and respiration to be calculated using a mass balance approach [@medhurst2006whole; @barton2010whole]. Here, we grew a single *Eucalyptus saligna* Sm. tree inside each of these 12 large, outdoor and sunlit WTCs for a period of 2 years. Each WTC can resolve net aboveground C uptake (canopy *A~n~* minus respiration of aboveground woody components), at high temporal resolution, while also controlling temperature and air humidity to track prevailing environmental conditions. Generally, measuring total canopy *A~n~* is difficult as variation in photosynthetic capacity exists within the canopy in response to the environment, requiring leaf measurements and models to upscale to the canopy [@ryan2010factors]. Combining continuous aboveground CO~2~ flux measurements with an evergreen *Eucalpytus* species, that grows throughout the year, enables tree C allocation to be tracked over long periods of time.

\
Previous findings in this experiment have shown that *Eucalyptus saligna* trees grown under eC~a~ treatments were smaller than ambient trees and that larger trees under ambient CO~2~ had a smaller reduction in canopy transpiration in drought conditions, via deeper rooting access to water resources [@duursma2011rooting]. The specific objectives of this study were to determine the response of biomass partitioning among foliage, aboveground woody components and roots of a native Australian tree species to changes in CO~2~ and altered water availability. Utilizing the unique WTC design we aimed to test how cumulative net aboveground C gain correlates to whole tree C mass increment, as a function of tree size. We then applied a mass balance approach to track the allocation of C above and belowground throughout the course of an 11 month period under the combined treatments of eC~a~ and drought.

\
Our hypotheses were:
\
(1) As C uptake and growth should be coordinated over long time periods, we expected both total leaf area and harvested tree C mass to correlate with cumulative total aboveground net canopy C uptake.

\
(2) At the end of the 2-year experiment we expected partitioning of C to harvested roots to increase under eC~a~. We also expected increases in partitioning to roots under drought treatments to alleviate water limitation.

\
(3) As increases in partitioning to root biomass were hypothesized, we expected TBCA to increase through time as cumulative tree C flux became affected by eC~a~ and drought.

#4.3  Methods

##4.3.1 Terminology
*Mass partitioning*:  the relative distribution of biomass between different tree tissue components such as leaves, branches, bole and roots.
\
*Carbon allocation*: the fraction of canopy photosynthesis distributed to different ecosystem components such as tissue biomass pools, respiratory C fluxes, non-structural carbohydrate storage pools or root C exudation.

##4.3.2 Whole tree chamber experiment
From April 2007 *Eucalyptus saligna* seedlings were grown in 12 whole-tree chambers (WTCs) at the Hawkesbury Forest Experiment in Richmond, Australia. One seedling per WTC (9 m high) was grown for 2 years and chamber conditions tracked outside air temperature and humidity. Each WTC was fitted with a root enclosure barrier that extended to the soil hard layer (1 m depth), separating WTC tree roots from neighboring trees. Roots were allowed to grow freely in the chamber soil volume and below 1 m. Full descriptions of the chamber design and operation are provided in Barton et al. [-@barton2010whole]. This multi-factor experimental design included CO~2~ × drought treatments with three WTC replicates in each of four treatments. Six chambers were kept at ambient CO~2~ of 380 ppm (aC~a~) and six were maintained at elevated CO~2~ of +240 ppm above ambient (eC~a~). Through October 2008 all trees were kept well-watered, with 10 mm of water every 3 days. Half of the chambers in each CO~2~ treatment were then subjected to a drought treatment by completely withholding water (dry) and the remaining six chambers were kept well-watered as an irrigated control (wet). The drought treatment lasted through mid-February 2009 when heavy rainfall ended the drought effect, despite the presence of a root enclosure [@duursma2011rooting].

##4.3.3 Aboveground chamber CO~2~ flux
Floors installed 45 cm above the soil surface, enclosing the main bole, permitted the chambers to function as cuvettes, excluding water and CO~2~ fluxes from the soil surface and allowed for whole tree fluxes of CO~2~ (and H~2~O) to be monitored once trees were ca. 3.5 m in height. This allowed high resolution CO~2~ flux data at 14 min intervals to be collected during the final eleven months of the experiment (from April 2008 to March 2009). The raw data were averaged over an hourly time-step and missing CO~2~ flux data were gap filled with SOLO (self-organizing linear output map) [see @abramowitz2005towards]. This self-fitting model predicted the flux as a function of photosynthetically active radiation, air temperature, vapor pressure deficit and day of year. In total, 27% of the observations were gap filled. This gap filling included a 5-6 week gap in late winter 2008 when WTC height was extended, as well as shorter periods were chamber access doors were open during regular use. For each WTC, cumulative 24 hour net aboveground C uptake ($\text{F}_{c,d}$, g C d^-1^) represented daily total canopy *A~n~* of each tree minus respiration of stems and branches. Then $\text{F}_{c,d}$ was summed over the flux monitoring period ($\text{F}_{c,T}$) to compare to tree C mass, leaf area and C allocation above and belowground.

##4.3.4 Harvested tree carbon mass
A final destructive harvest was completed in March 2009. The canopy of each tree was divided into five equal vertical layers, extending from the floor to the top and harvested. Dry biomass of leaves, branches and boles were measured for each layer and summed for each WTC. Root mass was obtained by excavating and sieving all soil inside each root exclusion barrier to the hard layer. Five root cores (10 cm diameter), sampled before the harvest, where collected from 0-70 cm in each chamber. Biomass from cores was added back to the standing crop total instead of scaling-up fine root biomass from cores to total chamber area. As a result, the fine root biomass pool is likely underestimated.

\
Carbon mass was assumed to be 48% of dry biomass for all non-leaf tissue components and this conversion was performed for all harvest and survey data (see below). This value represents the mean value of wood C of angiosperms from the Dyrad global wood C database, including measurements of stems, twigs, branches, bark, coarse roots and fine roots [@dryad_69sg2; @thomas2012treec]. Leaf and litter C mass was calculated by multiplying biomass by the WTC specific mean leaf C content (%). Leaf C content was determined from a sub-sample of final harvest dried and milled leaves analyzed using a Leco TruSpec Micro elemental analyzer (LECO corporation, MI, USA). Carbon mass fractions of leaves, boles+branches (stems) and roots were then calculated by dividing their respective total C mass by whole tree C mass at the end of the experiment.

\
Prior to the initiation of the experiment potted *Eucalyptus saligna* seedlings were harvested to develop relationships between above and belowground biomass (n=17, mean height of 217&pm;17 cm). These seedlings were grown in 25 l pots inside each WTC until the experiment started, using the same soil as each WTC, while chamber [CO~2~] treatment conditions were maintained. 

##4.3.5 Tree allometry surveys
Tree height was measured every 14 days and diameters were recorded monthly at regular intervals (30 cm) along the main bole and split stems. Bole diameters at 65 cm height were used as the starting reference diameter for each survey. Diameter and length for every branch, including forked branches, were surveyed seven times between April 2008 and March 2009. The first branch survey coincided with the installation of chamber floors and initiation of whole tree flux measurements. Branch diameter measurements were recorded at 5 cm from their individual insertion points. Leaf litter was collected from the chambers every two weeks, oven-dried and weighed.

##4.3.6 Bole carbon mass
During the final harvest, diameter measurements were recorded as described above and 1 cm wide cross sections were removed from the bole at equally spaced positions along the bole midpoint between the diameter measurement points. Wood density for each section was calculated by dividing the dry mass by the fresh volume separately for bark and wood. The mean total bole density for each tree ($\rho_{bo}$, g cm^-3^) was then calculated as the total density of bark and wood, weighted by the total diameter of each section. We assumed that $\rho_{bo}$ did not change through time.

\
For boles, individual volume units were constructed as concentric cylinders between each diameter measurement from base to tree top for each monthly survey. The tree top section was calculated as a cone with a tip radius of .001 cm. The volume below the starting reference diameter (65 cm) was calculated separately in order to interpolate taper into this section. Using the height of the tree and the standard diameter, the diameters at 30cm and base were estimated by extending the length of the pre-existing cone (from tree top to 65 cm). This resulted in two additional volume units. All volume units were summed,including forked stems, to calculate total bole volume. Bole mass was calculated as total volume multiplied by WTC-specific $\rho_{bo}$.

##4.3.7 Branch carbon mass
Measured dry mass, length and basal area of all harvested branches was used to determine the branch wood density ($\rho_{br}$) as well as a geometric shape factor [$\phi$, see @makela1997carbon] for each WTC tree by rearranging the equation:

\
(1)
$$M_{br} = L_{br}A_{br}\rho_{br}\phi_{br}$$

where $\text{M}_{br}$ is summed dry mass of all harvested branches, $\text{L}_{br}$ is summed branch length (cm), $\text{A}_{br}$ is summed branch basal area (cm^3^), $\rho_{br}$ represents the combined density of wood and bark (g cm^-3^) and $\phi$ corrects branch volume estimates to an intermediate shape between a cone and a cylinder [@makela1997carbon]. The ratio of measured  $\text{M}_{br}$ to $L_{br}A_{br}$ was used to generate a WTC-specific $\phi_{br}\rho_{br}$.

\
For each survey period, M~br~ was estimated by solving the above equation with $\text{L}_{br}$ and $\text{A}_{br}$ for individual branches with $\phi_{br}$$\rho_{br}$ specific to each WTC. We assumed that $\phi_{br}\rho_{br}$ did not change through time. Total dry branch mass at each survey point was the total mass of all individual branches.

##4.3.8 Leaf area and carbon mass
Total tree leaf area and dry mass were measured for each of the five canopy layers at the final tree harvest in March 2009. Specific leaf area (SLA, cm^2^ g^-1^) was calculated by dividing total projected one sided leaf area by dry leaf mass for each canopy layer. Mean SLA for each WTC tree was obtained by weighting SLA of each of the 5 layers by their foliage mass fraction. Estimates of standing leaf area were also obtained in April 2008 from leaf counts for each tree, multiplied by tree-specific mean leaf size (based on a sub-sample). 

\
Canopy leaf area was modeled on daily times steps, between April 2008 and March 2009, using the leaf count census and harvest leaf area estimates, along with height growth and litter fall rates. This was method was applied by Barton et al. [-@barton2012effects]. Briefly, leaf growth was assumed to coincide with height growth, so that no leaf growth occurred when height growth had ceased. This method assumes that total cumulative leaf area (i.e. standing leaf area plus that produced by litter fall) followed and allometric relationship with tree height such that: 

\
(2)
$$A_{L,T} = a~H^b$$
where $\text{A}_{L,T}$ is the total 'potential' leaf area (m^2^), a and b are tree specific coefficients and H is tree height (m). Then standing leaf area at time *t* are obtained from tree height at time *t* and cumulative litterfall:

\
(3)
$$A_{L} = a~H(t)^b - \int_{t=0}^t L(t)dt$$
where $\text{L(t)}$ is the litterfall (m^2^ t^-1^) rate at time *t*. Litter was assumed to be produced by all canopy layers. The daily leaf area contribution of litterfall is the difference between  $\text{A}_{L,T}$ and  $\text{A}_{L}$. The mean SLA for each harvested tree was multiplied by daily estimates leaf and litterfall area to calculate biomass. SLA for harvested trees was assumed to be constant over the entire flux measurement period.

##4.3.9 Tissue C allocation
Tissue specific C allocation represents the fraction of net canopy C uptake distributed to a given tissue, which determines the change in biomass of that tissue through time such that:

\
(4)
$$\frac{dM_{c}}{dt} = a_{c} NPP - \lambda_{c} M_{c}$$
where $\text{M}_{c}$ is the standing C mass of a component (g C), $\text{a}_{c}$ is the allocation to that component (0-1) and $\lambda_{c}$ is the component specific turnover (d^-1^).

\
Here, total C allocation to leaves and aboveground wood (branches + bole) could be estimated from the sums of tissue C mass, net aboveground C flux and tissue turnover for each day of the experiment such that:

\
(5)
$$M_{c,T} = \int a_{c} F_{c,d}dt - \int \lambda M_{c}dt$$
where $\text{M}_{c,T}$ is the total dry C mass of either leaves or wood and $\text{F}_{c,d}$ is the daily net aboveground C uptake (g C d^-1^). From equation 5, we estimated allocation by solving for $\text{a}_{c}$ (as all other components were measured). For example, C allocation to leaves ($\text{a}_{l}$) was determined by combining measurements of harvested dry C mass of leaves ($\text{M}_{l,T}$) with $\text{F}_{c,T}$ and total litterfall ($\text{L}_{l,T}$), giving:

\
(6)
$$M_{l,T} = a_{l} F_{c,T} - L_{l,t}$$
and then solving for leaf C allocation:

\
(7)
$$a_{l} = (M_{l,t} - L_{l,t}) / F_{c,t}$$

\
Allocation to aboveground wood C was estimated in the same manner with turnover measured as total dry C mass of branch litter collected across the experiment. For roots, only total belowground C allocation (TBCA) could be calculated (explained below) since root turnover was not measured.

##4.3.10 Total belowground carbon allocation
As the installation of chamber floors into each WTC separated the aboveground CO~2~ uptake from the soil CO~2~ efflux, TBCA at any time point *t* was calculated as:

\
(8)
$$TBCA_{t} = \int_{0}^t F_{c,d}dt - C_{ab}$$
where $\text{C}_{ab}$ is the aboveground standing crop C mass (g C) of stems, branches, leaves and total leaf litterfall. As the final standing crop of root biomass was known, TBCA could be further broken down into the total C mass of roots ($\text{C}_{r,T}$) and the residual belowground C flux ($\text{F}_{c,r}$). The residual belowground C flux includes root and microbial respiration, root turnover, root exudation and any unaccounted for root C mass. The use of aboveground allometry to interpolate $\text{C}_{ab}$ through time combined with measured daily $\text{F}_{c,d}$ allowed TBCA to be estimated on daily time steps over the final eleven months of the experiment while cumulative $\text{F}_{c,r}$ was calculated at the final harvest.

##4.3.11  Mass balance relationships between $\text{F}_{c,d}$ and carbon allocation.
The cumulative sum of $\text{F}_{c,d}$ for each WTC, at any given time point, represented the running total of net C uptake since the chamber floors were installed. Daily allocation of C to boles and branches was estimated by linear interpolation between 14-day survey measurements and the final harvest, starting at the first branch survey (April 2008). These daily estimates of leaf and litter C were added to bole and branch C mass to estimate $\text{C}_{ab}$ on any given day. The contribution of each aboveground component to the cumulative sum of $\text{F}_{c,d}$ was then tracked from April 2008 to March 2009. The initial estimated C mass of each aboveground component and $\text{F}_{c,d~}$ on the day when chamber floors were installed was subtracted from all respective daily values so mass balance could be tracked with a 0 starting value. This allowed cumultaive estimates of TBCA to be generated across the final 11 months of the experiment. Additionally, the significant log-linear relationship between above and belowground mass of both harvested trees and potted seedlings (R^2^ = 0.98, Figure `r rootshootpred`) was used to estimate $\text{C}_{r,T}$ from $\text{C}_{ab}$ on the last day of the 11 month period. The major carbon allocation pathways and subsequent above- and belowground C pools measured in each WTC are further depicted in Figure `r conceptfig`.

\
```{r concept, echo=FALSE, error=TRUE,fig.width=8, fig.height=6,dev="png", dpi=600, dev.args=list(type="cairo")}
library(png)
concept <- readPNG("manuscript/conceptfig4.png")
par(mar=c(0,0,0,0))
plot(c(0,1),c(0,1), type="n", axes=F, xlab="", ylab="")
rasterImage(concept, xleft = 0, ybottom = 0, xright = 1, ytop =1)

```
\
**Figure 4.1**. Conceptual diagram depicting the major components of C flow among plant components including; uptake via photosynthesis, allocation to component tissues, tissue respiration and root exudation. Mass pools of leaves, wood and roots include tissue turnover. Net aboveground C uptake ($\text{F}_{c}$), shown in the shaded box, represents the flux of C measured within each WTC.  With the WTC experimental design, total belowground C allocation (TBCA) is measured as the residual between $\text{F}_{c,T}$ and total aboveground C mass.  

##4.3.12 Data analysis
Differences in experimental parameters to the interaction of CO~2~ and drought treatments at the final harvest were analysed as a completely randomized experimental design with factorial treatment combinations using two-way ANOVA in R [@RDevelopmentCoreTeam2011]. Tukey's post-hoc tests were performed in conjunction with ANOVA to determine which specific paired comparisons among climate change treatments were different. Significance level was set at an alpha of 0.05 and findings between 0.05 and 0.10 were considered marginally significant.

#4.4  Results

##4.4.1 Total aboveground carbon flux, whole tree C mass and leaf area
Both whole tree C and $\text{C}_{ab}$ from the final harvest were reduced in eC~a~ treatments by 32 % (both P < 0.03). Over the entire 11 month measured chamber flux period the summed aboveground C uptake ($\text{F}_{c,T}$) was significantly reduced by 30.5 % in eC~a~ treatments (P = 0.043), while no effects of the drought treatment were detected (Table `r wtc1tab1`). $\text{F}_{c,T}$ was positively correlated with estimates of both whole tree C (R^2^ = 0.74, Figure `r fluxtreefig`a) and $\text{C}_{ab}$ (R^2^ = 0.69, Figure `r fluxtreefig`b) over the same time period. Whole tree C mass increment estimated during the chamber flux period represented ca. 75 % of total harvested tree C mass. As the majority of biomass production occurred during this period, the allometric estimates of whole tree C were used for comparison to $\text{F}_{c,T}$.

\
**Table 4.1**. Final harvest C mass of above and belowground tissues, cumulative aboveground tree C uptake ($\text{F}_{c,T}$) and specific leaf area (SLA). Each value represents the mean (&pm; 1 standard error) for each treatment combination. Units for C mass and $\text{F}_{c,T}$ are g C, while SLA are cm^2^ g^-1^. For each variable, different letters represent significant differences between treatments from the overall model which includes CO~2~ * drought interactions. P values represent overall differences of CO~2~ or drought main effects and the CO~2~ * drought interaction.
```{r table, echo=FALSE}
table2 <- rbind(table, c("CO~2~ effect (P)",  Pco2_2$x))
table3 <- rbind(table2, c("Drought effect (P)",  Ph2o_2$x))
table4 <- rbind(table3, c("CO~2~ * Drought (P)",  Pinter_2$x))

dust(table4) %>%
    sprinkle(rows= 1, part="head",bold=TRUE)%>%
    sprinkle_colnames(Tree.C.flux = "F~c,T~") %>%
    sprinkle_print_method("markdown")
```

\
```{r fluxmass, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/Cflux_treecarbon_figure3.R")
```
\
**Figure 4.2**. Whole tree C mass as a function of cumulative aboveground C flux for each WTC tree. Values of cumulative aboveground net C flux were measured over the final eleven months of the experiment. Whole tree C mass represents the sum of bole, branch, leaf and root C mass from allometric estimates over the same time period. The dotted line is the 1:1 relationship and the solid line represents the significant overall linear model fit from the equation y = 0.56x + 878.2 (R^2^ = 0.86).

\
Leaf area at the final harvest was significantly reduced by by 31.3 % under eC~a~ (p < 0.001) and this pattern was observed across the final eleven months of the experiment (Figure `r lafig`). Specific leaf area was reduced by 10.9 % in eC~a~ treatments (P = 0.053), and by 8.9 % in drought treatments (P = 0.089, Table `r wtc1tab1`). Overall, $\text{F}_{c,T}$ was positively correlated with mean leaf area (P < 0.001, Figure `r fluxlafig`). Intercepts and slopes between separate linear regressions of $\text{F}_{c,T}$ and mean leaf area for aC~a~ and eC~a~ treatments were not different, however, there was negligible overlap of data between treatments. Thus, we were unable to determine if reductions in $\text{F}_{c,T}$ in eC~a~ treatments were a function of lower mean leaf area or shifts in *A~n~* and tissue respiration rates.

\
```{r leafarea, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/leafarea_figure.R")
```
\
**Figure 4.3**. Estimated canopy leaf area for each WTC tree over the final eleven months of the experiment (April 2008 to March 2009). Estimates are based on height growth, litterfall rates and two leaf area estimates following Barton et al. [-@barton2012effects]. Color and line type distinguish the treatment combination for each WTC. 

```{r fluxla, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/Cflux_leafarea_regressions.R")
```
\
**Figure 4.4**. Treatment means of cumulative aboveground C flux as a function of mean daily canopy leaf area over the final eleven months of the experiment. The solid line represents the significant overall linear model fit (R^2^ = 0.77) from the equation: y = 611.9x + 2791.2. Separate 95% confidence intervals are shown for linear regression between $\text{F}_{c,T}$ and mean leaf area for aC~a~ and eC~a~ treatments.

##4.4.2 Harvested tissue carbon mass and biomass partitioning
At the end of this two year experiment, harvested C mass of tissue components was affected in eC~a~ but not drought treatments (Table `r wtc1tab1`). Aboveground wood C mass was reduced by 37 % in eC~a~ treatments (P = 0.015), driven mostly by eC~a~ effects on bole wood. Neither standing crop leaf C mass or total litterfall C mass over the study period differed between CO~2~ treatments. Total root C mass was  reduced by 29% in eC~a~ treatments (P = 0.091).

\
Leaf mass fraction (LMF) increased by 15.0 % in eC~a~ treatments (P = 0.011) but was not affected by the drought treatment. Leaf mass fraction was negatively correlated with whole tree C mass (P= 0.007, Figure `r massfracfig`a). Stem mass fraction (SMF) was marginally reduced by 6.0 % under eC~a~ (P = 0.077), with no effect of the drought treatment detected. Stem mass fraction had a weak positive correlation with whole tree C mass (P = 0.08, Figure `r massfracfig`c). Root mass fraction (RMF) was not affected by either treatment and was not correlated to whole tree C mass (Figure `r massfracfig`e).

\
```{r c_allocation, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/fractionallocation_fig2.R")
```
\
**Figure 4.5**. Treatment means of C mass fractions of leaves (a), stems (branches+boles) (c) and roots (e) as a function of tree size, via whole tree C mass. Treatment means of C allocation to leaves (b) and stems (d) as a function of cumulative aboveground net C flux. Root C allocation could not be estimated as root turnover was not known. Values for C mass fractions are calculated from final harvest biomass totals. Values for C allocation are estimated from cumulative total aboveground net C flux over the final eleven months of the experiment. Solid lines represent overall linear model fit for leaf, stem and root mass fractions (R^2^ = 0.53, 0.26 and 0.01, respectively), as well as leaf and stem C allocation (R^2^ = 0.39, 0.01, respectively).

##4.4.3 Aboveground carbon allocation
Treatment effects on tissue C allocation were determined from C mass estimates obtained from allometry over the final eleven months of the experiment and $\text{F}_{c,T}$ over the same time period. Total C allocation to leaves increased by 28% in eC~a~ treatments (P = 0.052), with no effect of the drought treatment detected. Leaf C allocation was negatively correlated with $\text{F}_{c,T}$ (P = 0.031, Figure `r massfracfig`b). Alternatively, C allocation to aboveground wood was not affected by either treatment and was not correlated to whole tree C (Figure `r massfracfig`d).

##4.4.4 Belowground carbon allocation
Across all treatment combinations, the total C mass of boles, branches, leaves and roots produced through the course of the measured flux measurement period was on average 61.0&pm;0.02 % of $\text{F}_{c,T}$ (Figure `r additivemassfig`). As mass balance must be achieved, TBCA and the residual belowground C flux ($\text{F}_{c,r}$) were estimated from Figure `r additivemassfig` as residuals between $\text{F}_{c,T}$ and whole tree mass excluding and including estimates of roots over the flux measurement period, respectively. Total belowground C allocation was on average 49.9&pm;0.02 of $\text{F}_{c,T}$ and ranged from 46.1 to 54.9 % across treatment combinations. Across a large range in tree size among the treatment combinations and replicate WTCs, similar patterns were detected for each tree (Figure `r additivechamfig`). Neither cumulative TBCA nor $\text{F}_{c,r}$ were affected by CO~2~ or drought treatments (Figure `r belowbarfig`). The time course of cumulative daily TBCA and $\text{F}_{c,T}$ were positively correlated over the final 11 months of the experiment (R^2^ = 0.78, P < 0.001) and the proportion of C allocated belowground was relatively stable through time and between treatments (Figure `r tbcafluxfig`).

\
```{r c_cumulative, fig.width=7, fig.height=10, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/cumulativecarbon_figure.R")
```
\
**Figure 4.6**. Cumulative aboveground net C flux and additive C allocation to individual tree components from 15 April 2008 to 16 March 2009. Each panel represents mean values for each treatment combination (n=3). Both aboveground net C flux and tissue C allocation where set to 0 on 15 April 2008 in order to track the allocation of C in daily time steps. Root C mass, predicted from the logarithmic relationship between above and belowground mass partitioning of pre-planting seedlings and harvested trees, is shown on the last date.

```{r belowbar, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/tbca_figure2.R")
```
\
**Figure 4.7**. Treatment means &pm; 1 standard error of cumulative aboveground net C flux, TBCA, and the residual belowground C flux ($\text{F}_{c,r}$). Values of cumulative aboveground net C flux were measured over the final eleven months of the experiment. Values for TBCA are the residual between the cumulative C flux and total C mass aboveground estimated from allometric surveys over the same time period. Values for $\text{F}_{c,r}$ were calculated as the residual between TBCA and root C mass predicted on the last date of the eleven month period.  

```{r tbcaflux, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/Cflux_TBCA_figure.R")
```
\
**Figure 4.8**. Total belowground C allocation as a function of cumulative aboveground net C flux across the final eleven months of the experiment. Carbon mass aboveground was estimated from allometric surveys, interpolated on a daily time scale and then subtracted from the aboveground net C flux to quantify TBCA. Individual lines represent treatment means, with color and line type distinguishing treatment 

#4.5  Discussion
A whole-tree chamber experiment provided a unique opportunity to study the C balance of *Eucalyptus* trees. We found that biomass partitioning and C allocation of component tissues were differentially affected by eC~a~. Despite previous findings of negative impacts of drought on leaf and canopy physiology in this study [see @duursma2011rooting; @crous2012light], minimal effects of a four month drought were detected on total tree C flux, biomass partitioning and tissue C allocation. Using a novel methodological framework, we show that TBCA may be less sensitive to climate change factors than previously assumed. As reliable estimates of TBCA are notoriously hard to obtain, we provide essential empirical data that can be compared to model predictions where C allocation is represented.

##4.5.1 Relationships between tree C flux, leaf area and tree C mass
A novel aspect of this study was the ability to measure whole tree C fluxes directly and compare these fluxes to observed patterns in leaf area and growth. Tree C uptake and growth were strongly coordinated across this two year experiment. The net C uptake of plants should be a function of the canopy leaf area and light interception [@wilson1965stand; @monsi2005factor] and correlate to canopy assimilation and tree productivity [@lindroth2008leaf; @mccarthy2006canopy; @waring1983estimating]. Estimates of tree canopy C flux, however, are limited by simple upscaling of single leaf measurements [@amthor1994scaling], oversimplification of big leaf models [@de1997simple] or parameterization of more complex models with assumptions of canopy behavior [@leuning1995leaf]. We found that leaf area was consistently reduced in eC~a~ treatments, likely leading to reductions in both tree C uptake and whole tree C mass of near identical magnitudes (ca. 30 %).  

\
A shortage of empirically measured whole tree C fluxes also makes inferring relationships with biomass or C allocation difficult. Biomass and C fluxes have been found to be poorly related in forest ecosystems due to difficulty in accounting for C retention of different tissues [@litton2007carbon]. This partial accounting of C likely inhibits the ability of many studies to precisely test the coordination between canopy *A~n~* and growth. The advantage of the WTC approach is the ability to compare cumulative whole tree C fluxes to absolute biomass production over long time periods. Here, we show empirically measured aboveground tree C uptake (F~c~) was strongly correlated to tree biomass production across a 2.5 fold size range in *Eucalyptus* trees. 

##4.5.2 Responses of biomass partitioning and C allocation to climate change
We first used final harvest biomass to determine patterns of biomass partitioning to leaves, stems and roots. We then combined cumulative tree C fluxes with tissue biomass production and turnover to measure C allocation to stems, leaves and TBCA, via mass balance. This approach allowed us to evaluate the impacts of climate change treatments on tree growth through potential shifts in tissue biomass production or C allocation. This is because there are many possible fates for C assimilates beyond just the production of plant biomass [@korner2005carbon]. Changes in C allocation encompass effects of tissue turnover, the storage and use of carbohydrates and root exudation to stimulate microbial activity, with each representing significant tree or ecosystem responses to environmental change. Thus, patterns in biomass partitioning and C allocation may not be consistent with respect to the tissue in question, which contributes to the current uncertainty in modelling tree growth responses to interacting climate change factors.

\
We found that stem C mass was reduced in eC~a~ treatments. Opposite responses of stem growth under eC~a~ have been found across different forested FACE experiments, including no effect in a mixed deciduous forest at WEB-FACE [@korner2005carbon] and a positive enhancement in a loblolly pine forest at duke FACE [@delucia2005contrasting]. It is possible that observed patterns in stem C mass were related to allometric trajectories as a function of plant size [@tjoelker1998temperature; @muller2000effect] more than direct effects of eC~a~ on stem biomass production. Stem mass fractions (SMF) were found to increase with total plant size and were marginally reduced in eC~a~ treatments. Carbon allocation to stems was unaffected in eC~a~ treatments, however, inferring that patterns in SMF were a consequence of size-dependent relationships between larger aC~a~ trees compared to smaller eC~a~ trees. Trees in this experiment followed commonly observed developmental patterns in biomass partitioning, with increases in SMF and decreases in LMF as tree became larger [@poorter2015does]. Thus, it is likely that eC~a~ treatments negatively affected other tree processes which first decreased overall tree size.  

\
Contrary to expectation, we found that both LMF and C allocation to leaves increased in eC~a~ treatments independent of tree size effects. As leaf production and turnover were not subsequently affected in the smaller eC~a~ trees, it is likely that changes in other physiological processes were necessary to explain observed increases leaf C allocation. Previously reported increases in leaf respiration under eC~a~ treatments [@crous2012light] are intrinsically included in the measurement of $\text{F}_{c,d}$, thus observed increases in leaf C allocation in terms of leaf biomass production are independent of shifts in respiration. Decreases in SLA were detected in WTC trees under eC~a~ treatments, which is often found across eC~a~ enrichment studies [@ainsworth2005have; @wang2012meta; @yin2002responses]. Concentrations of leaf non-structural carbohydrates (TNC) often increase under eC~a~ [@Roden01071996; @Picon1997; @poorter1997; @loewe2000mycorrhiza; @walter2005] and are often associated with subsequent decreases in SLA in trees [@korner2005carbon; @barron2005growth]. Here, increased C allocation to leaves may have resulted in increased leaf TNC to fulfill increased canopy respiratory demands or meet sink demands of other tissues. Taken together, results for aboveground tissues highlight the importance of separating impacts on measured biomass from those of total C allocation associated with growth  when evaluating tree responses to climate change.

##4.5.3 TBCA response to climate change in a single-tree ecosystem
Despite increased attention to the effects of climate change on belowground processes, the difficulty in measuring TBCA currently hinders our ability to make well-founded empirical conclusions. One of our specific objectives was to use a novel method to calculate TBCA to test the hypothesis that TBCA was enhanced in eC~a~ treatments and then to evaluate potential shifts in TBCA across shorter times scales. For example, changes in TBCA to eC~a~ or drought could occur as sustained or pulsed responses through time. Enhancement of TBCA has been reported across forested FACE experiments [@palmroth2006aboveground] but the single-tree ecosystem design of the WTC allowed us to evaluate the effects of climate change factors without the inherent environmental complexity of a forest community. The unique design of the WTC allowed us to track TBCA as a cumulative total and across daily time steps, both of which can be used to validate and constrain models where C allocation is represented.

\
With high resolution flux data and reliable estimates of aboveground dry mass production we show that TBCA was not affected by eC~a~ or drought treatments over the final eleven months of the experiment. Contrary to expectation, we detected minimal effects of eC~a~ or drought treatments on root biomass partitioning, although it was not possible to differentiate fine and coarse roots production and turnover. Although these findings disagree with TBCA results from forested FACE experiments [see @palmroth2006aboveground], comparisons between single-tree studies with evidence from forest ecosystem experiments should be made with caution. Nevertheless, we show that TBCA in *Eucalyptus* trees may be less sensitive to climate change factors than expected over a ~1 year period. However, a lack of cumulative change in TBCA does not infer that belowground processes were not affected by either treatment. In trees under drought stress, TBCA might increase with higher allocation to root systems to alleviate water stress [@poorter2012biomass], which could by offset increased root mortality and turnover [@marshall1986drought; @meier2008belowground], reduced root exudation [@iversen2014terrestrial] or reduced C demand via decreases in root respiration rates [@burton1998drought]. Alternatively, the lack of belowground competition for soil mineral resources in this single tree ecosystem might have delayed enhancement of TBCA to eC~a~ treatments, such as increased root production and exudation.

\
With an estimation of daily aboveground C mass gain and measured cumulative whole tree C uptake we were able to uniquely track dynamic short term effects of eC~a~ or drought on TBCA. Across daily time steps, we observed a relatively stable fraction of total tree C flux distributed to TBCA over a period of eleven months. The ability to calculate TBCA as a simple residual between measured aboveground processes gives us reliable estimates of the absolute amount of C distributed belowground each day, which appears to be insensitive to sustained exposure to eC~a~ and a four month drought. Similar to Palmroth et al. [-@palmroth2006aboveground] we cannot quantify allocation to specific belowground pools, but our approach with the WTC design does not have to make assumptions about C residence time in any tissue or soil component. As a result, the lack of a cumulative response of TBCA raises questions about the regularity of belowground responses to climate change factors often reported. Our results confirm the need for more reliable estimates of TBCA in future studies, which are crucial for predicting forest responses to climate change. 

##4.5.4 Conclusions
Here we use novel aspects of the WTC experimental facility to show that whole tree C flux and tree growth were highly correlated, while patterns in biomass partitioning alone were insufficient to explain eC~a~ effects on tree growth. With individual *Eucalyptus saligna* trees we show different responses of above and belowground C allocation to eC~a~ treatments, which has important implications for how C allocation should be represented in process-based forest models. As empirical measurements of belowground processes are still difficult to obtain, models may have to assume that responses of aboveground tissues to global change represent those of belowground tissues [@giardina2005response]. As a result, continued empirical measurements to define C allocation patterns constrained by functional relationships with biomass production are needed to reduce uncertainty and improve model predictions [@de2014does]. Continuing to apply novel approaches to better evaluate TBCA and empirically measure whole tree C fluxes, such as the WTC experiment, are the way forward in addressing questions regarding the fate of assimilated C under global climate change.

#4.6  Supporting Information {.unnumbered}

```{r rootshoot, fig.width=7, fig.height=7, message=FALSE,echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/root_predict.R")
```
\
**Figure 4.S1**. Root mass as a function of shoot mass in *Eucalyptus saligna* for potted seedlings harvested before planting of WTC trees (n=17) and WTC trees harvested after 2 years (n=12). Potted seedlings were grown in 25 l pots inside each WTC, while chamber [CO~2~] treatments conditions were maintained. The solid line represents the significant log-log model fit (R^2^ = 0.98) from the equation: log(x) = 0.77(log(y)) + 0.43.

```{r chambers, fig.width=8, fig.height=11, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/cumulativecarbon_cham_figure2.R")
```
\
**Figure 4.S2**. Cumulative aboveground net C flux and additive C allocation of individual tree components from 2008-4-15 and 2009-3-16. Panels represent each individual WTC. Both aboveground net C flux and tissue C allocation where set to 0 on 2008-4-15 in order to track the allocation of C in daily time steps. Total root C mass, predicted from the log relationship between above and belowground mass partitioning of pre-planting seedlings and harvested trees, is shown on the last date.


# References {.unnumbered}